SI_Appendix_Datasets

### 2024-06-16 R/SI_Appendix_Datasets.Rmd
plots: examine GxE in QTL and mash SNP overlaps

Get `qtlplot` object by running code starting at about line 150 (for Dataset 7). Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).

@@

  mash_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      dplyr::mutate(value = row_number()) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value) |> 
  ## TODO: FIX list/matrix column here
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, 
                    sep = "_", extra = "merge", convert = TRUE) %>%
    dplyr::arrange(.data$Chr, .data$Pos)

@@

    if(i == 1){
    outputdf <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
                       QTL_Pos_hi = qtl$POS_hi[i], 
                       QTL_phenotype = qtl$lodcolumn[i], 
                       QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
                       mash_log10BayesFactor = qtl_sub$log10BayesFactor, mash_subpop = subpop, 
                       mash_phenotype = phe)
    } else {
      outputdf1 <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
                          QTL_Pos_hi = qtl$POS_hi[i], 
                          QTL_phenotype = qtl$lodcolumn[i], 
                          QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
                          mash_log10BayesFactor = qtl_sub$log10BayesFactor,
                          mash_subpop = subpop, 
                          mash_phenotype = phe)
      outputdf <- outputdf %>%
        full_join(outputdf1)
    }
  }
  outputdf <- outputdf |> 
    mutate(
      Marker_in_mash_1pct_tail = 
        case_when(mash_log10BayesFactor >= onepcttail ~ TRUE,
                  mash_log10BayesFactor < onepcttail ~ FALSE,
                  TRUE ~ NA)
    )
  return(outputdf)
}

@@

>>>>

Get `qtlplot` object by running code starting at about line 233 (for Dataset 7). Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).

@@

  mash_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      dplyr::mutate(value = row_number()) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      #dplyr::select(-.data$value) |> 
  ## TODO: FIX list/matrix column here
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, 
                    sep = "_", extra = "merge", convert = TRUE) %>%
    dplyr::arrange(.data$Chr, .data$Pos)

@@

    if(i == 1){
    outputdf <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
                       QTL_Pos_hi = qtl$POS_hi[i], 
                       QTL_phenotype = qtl$lodcolumn[i], 
                       QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
                       mash_idx = qtl_sub$value,
                       mash_log10BayesFactor = qtl_sub$log10BayesFactor, mash_subpop = subpop, 
                       mash_phenotype = phe)
    } else {
      outputdf1 <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
                          QTL_Pos_hi = qtl$POS_hi[i], 
                          QTL_phenotype = qtl$lodcolumn[i], 
                          QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
                          mash_idx = qtl_sub$value,
                          mash_log10BayesFactor = qtl_sub$log10BayesFactor,
                          mash_subpop = subpop, 
                          mash_phenotype = phe)
      outputdf <- outputdf %>%
        full_join(outputdf1)
    }
  }
  outputdf <- outputdf |> 
    mutate(
      Marker_in_mash_1pct_tail = 
        case_when(mash_log10BayesFactor >= onepcttail ~ TRUE,
                  mash_log10BayesFactor < onepcttail ~ FALSE,
                  TRUE ~ NA)
    )
  return(outputdf)
}

@@

### Exploratory effect plots for mash QTL overlaps

2N 59.5Mb: biggest effect at TX2, mostly opposite to other gardens or other gardens have NSE
2N 60.9Mb: biggest effect at TX2, AP between TX1 and TX2
2 Gulf effects I would say look nothing like that.
```{r}
overlap <- dataset_8 %>%
  select(mash_idx, mash_subpop, everything()) |> 
  distinct(mash_idx, .keep_all = TRUE)
overlap
```

```{r}
refactor_conditions <- function(phe, subpop, idx) {
  if (phe == "Flowering Date" & subpop == "Both") {
    plot <- mash_plot_effects(m_gam_fl_hyp, i = idx)
    plot$effect_df$Conditions <- 
      factor(plot$effect_df$Conditions, 
             levels = c(paste0("Gulf_and_Midwest_FL50_", c("TX1", "TX2", "TX3",
                                                           "OK", "MO", "NE",
                                                           "MI", "SD"))))
  } else if (phe == "Flowering Date" & subpop == "Gulf") {
    plot <- mash_plot_effects(m_gulf_fl_hyp, i = idx)
    plot$effect_df$Conditions <- 
      factor(plot$effect_df$Conditions, 
             levels = c(paste0("Gulf_FL50_", c("TX1", "TX2", "TX3",
                                               "OK", "MO", "NE",
                                               "MI", "SD"))))
  } else if (phe == "Flowering Date" & subpop == "Midwest") {
    plot <- mash_plot_effects(m_midw_fl_hyp, i = idx)
    plot$effect_df$Conditions <- 
      factor(plot$effect_df$Conditions, 
             levels = c(paste0("Midwest_FL50_", c("TX1", "TX2", "TX3",
                                               "MO", "M", "SD"))))
    } else if (phe == "Greenup Date" & subpop == "Both") {
    plot <- mash_plot_effects(m_gam_gr_hyp, i = idx)
    plot$effect_df$Conditions <- 
      factor(plot$effect_df$Conditions, 
             levels = c(paste0("Gulf_and_Midwest_GR50_", c("TX1", "TX2", "TX3",
                                                           "OK", "MO", "NE",
                                                           "MI", "SD"))))
  } else if (phe == "Greenup Date" & subpop == "Gulf") {
    plot <- mash_plot_effects(m_gulf_gr_hyp, i = idx)
    plot$effect_df$Conditions <- 
      factor(plot$effect_df$Conditions, 
             levels = c(paste0("Gulf_GR50_", c("TX1", "TX2", "TX3",
                                               "OK", "MO", "NE",
                                               "MI", "SD"))))
  } else if (phe == "Greenup Date" & subpop == "Midwest") {
    plot <- mash_plot_effects(m_midw_gr_hyp, i = idx)
    plot$effect_df$Conditions <- 
      factor(plot$effect_df$Conditions, 
             levels = c(paste0("Midwest_GR50_", c("TX1", "TX2", "TX3", "OK",
                                               "MO", "NE", "MI", "SD"))))
  } else stop("not built in yet")
  plot$ggobject <- ggplot(data = plot$effect_df) + 
      geom_point(mapping = aes(x = as.factor(.data$Conditions), 
                               y = .data$mn)) + 
      switchgrassGWAS::theme_oeco + 
      geom_errorbar(mapping = aes(ymin = .data$mn - 
                                    .data$se, ymax = .data$mn + .data$se, 
                                  x = .data$Conditions), 
                    width = 0.3) + geom_hline(yintercept = 0, lty = 2) + 
        labs(x = "Conditions", y = "Effect Size") + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      coord_flip() +
      ggtitle(plot$marker)
  return(plot)
}
```

```{r}
gr_qtl_overlap <- overlap |> 
  mutate(row_n = row_number(),
         .before = 1) |> 
  filter(mash_phenotype == "Greenup Date") |> 
  arrange(Chr, desc(mash_log10BayesFactor)) |> 
  select(row_n, mash_subpop, mash_log10BayesFactor, Chr, QTL_Pos_lo,
         Marker_in_mash_1pct_tail, everything())

fl_qtl_overlap <- overlap |> 
  mutate(row_n = row_number(),
         .before = 1) |> 
  filter(mash_phenotype == "Flowering Date") |> 
  arrange(Chr, desc(mash_log10BayesFactor)) |> 
  select(row_n, mash_subpop, mash_log10BayesFactor, Chr, QTL_Pos_lo,
         Marker_in_mash_1pct_tail, everything())
```
row_n 117: No match to QTL effects, just large Midwest mash effect at MO

9N QTL: (-) effects at OK & MI.

For greenup, all but 3 significant mash SNPs (in the Gulf subpop) have just a single significant effect at MO. No QTL are like this, the only strong effect at MO is on Chr3K in the QTL. The 3 Gulf QTL don't really look like the 9N QTL either.

```{r}
i = 142 # 152
for (i in 1:nrow(gr_qtl_overlap)) {
  i <- gr_qtl_overlap$row_n[i]
  replot <- refactor_conditions(phe = overlap$mash_phenotype[i],
                    subpop = overlap$mash_subpop[i],
                    idx = overlap$mash_idx[i])
  save_plot(filename = here("manuscript", "revision2", "intermediate_files", "mash_effect_plots", "greenup", paste0(str_replace_all(replot$marker, pattern = " ", "_"), "_effect.png")), plot = replot$ggobject)
}

for (i in 1:nrow(fl_qtl_overlap)) {
  i <- fl_qtl_overlap$row_n[i]
  replot <- refactor_conditions(phe = overlap$mash_phenotype[i],
                    subpop = overlap$mash_subpop[i],
                    idx = overlap$mash_idx[i])
  save_plot(filename = here("manuscript", "revision2", "intermediate_files", "mash_effect_plots", "flowering", paste0(str_replace_all(replot$marker, pattern = " ", "_"), "_effect.png")), plot = replot$ggobject)
}
```
